---
title: "Statistical Learning Project"
author: " Filippo Santin, Gurjeet Singh, Francesca Zen"
date: "18/5/2021"
output:
  pdf_document: default
  html_document: default
---

# Exploiting the Dataset

In the following report we present the analysis computed on the stroke dataset provided by kaggle ^[https://www.kaggle.com/fedesoriano/stroke-prediction-dataset].
```{r}
stroke_data <- read.csv('healthcare-dataset-stroke-data.csv')
attach(stroke_data)
```

It is composed of 5,110 entries with a total of 12 columns: id, gender, age, hypertension, heart_diseas, ever_married, work_type, Residence_type, avg_glucose_level, bmi, smokin_status, stroke.
Our aim is to see how the features were related to each other in order to predict which individual is more probable to get a stroke.
The preliminary part of the analysis focuses on the study of the dataset, we remove the column which identify the singular individuals 
```{r}
stroke_data<-stroke_data[,-1]
```

and transformed the categorical variables into factors:

```{r}
stroke_data$gender<- as.factor(gender)
stroke_data$ever_married<-as.factor(ever_married)
stroke_data$work_type<-as.factor(work_type)
stroke_data$Residence_type<-as.factor(Residence_type)
stroke_data$smoking_status<-as.factor(smoking_status)
```

We discovered that the variable bmi was not numeric because of the presence of a string "N/A" which identifies the lack of the value, and so we transform it using the command `stroke_data$bmi <- as.numeric(bmi)`. We then removed those NA values:`stroke_data<- na.omit(stroke_data)`.
We ended up having  4,909 entries and 11 total columns.
Here we give a quick overview of the main information about the dataset:

```{r}
summary(stroke_data)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
